\documentclass[]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\pagenumbering{gobble}

\begin{document}

\section {Extended fields}

Let $F$ be a totally ordered field.
Let $F_\infty = F \cup \{ \bot, \top \}$.
We define operations $+$ and $\cdot$ on $F_\infty$
so that
(for all $a \in F_\infty$ and for all $p \in F, p > 0$)
from top down:
\begin{align*}
	\bot <& \;a\, < \top \\
	\bot + a =& \;\bot = a + \bot \text{\qquad(includes $\bot + \top = \bot$)} \\
	\top + a =& \;\top = a + \top \\
	\bot \cdot\, p =& \;\bot = p \,\cdot \bot \\
	\top \cdot\, p =& \;\top = p \,\cdot \top \\
	\bot \cdot\, 0 =& \;\bot = 0 \,\cdot \bot \text{\qquad\;(the ``weird'' rule)} \\
	\top \cdot\, 0 =& \;\,0\, = 0 \,\cdot \top
\end{align*}
All other cases of all operations and relations
preserve their behavior from $F$.
We keep the product between negative numbers and
$\{ \bot, \top \}$ undefined.

\section{Farkas-like conjecture}

Let $A \in F_\infty^{m \times n}$ and $b \in F^m$.
Exactly one of the following is true:
\begin{itemize}
	\item $\exists x \in F^n$ such that
	$0 \leqq x$ and $A x \leqq b$
	\item $\exists y \in F^m$ such that
	$0 \leqq y$ and $(-A^T) y \leqq 0$ and $b^T y < 0$
\end{itemize}

\section{Counterexample}

$$
A =
\begin{pmatrix}
	\bot & \top\\
	\top & \bot
\end{pmatrix}
\qquad \qquad
b =
\begin{pmatrix}
	-1 \\
	-1
\end{pmatrix}
$$
Both are true.

\section{Remark}

For all other versions of Farkas lemma
that I tried to generalize to $F_\infty$
similar counterexample still applies.
However, they might work if it was forbidden
for a row of $A$ to contain both $\bot$ and $\top$
and alike for a column.

The problem is when the conversion to the finite
version requires us to delete both rows and columns,
the rows must be deleted first both in the primar
and in the dual, and so, if deleting a row stops
deleting a column from triggering, it leads to a
different result in the primar than in the dual
because of the matrix transposition.

\newpage
\section{New attempt}

Let $A \in F_\infty^{m \times n}$ and $b \in F_\infty^m$.
Assuming that no row and no column of $A$ contains
both $\bot$ and $\top$ elements and that
$A$ does not have $\bot$ on any row where $b$ has $\bot$,
exactly one of the following is true:
\begin{itemize}
	\item $\exists x \in F^n$ such that
	$0 \leqq x$ and $A x \leqq b$
	\item $\exists y \in F^m$ such that
	$0 \leqq y$ and $(-A^T) y \leqq 0$ and $b^T y < 0$
\end{itemize}

\section{Proof sketch}

\begin{itemize}
\item $ A \in F_\infty^{I \times J} $
\item $ b \in F_\infty^{I} $
\item if $\bot \in b$ then easy; assume $\bot \notin b$
\item $ I' := \{ i \in I ~|~ b_i \neq \top \wedge \bot \notin A_{i, \star}  \} $
\item $ J' := \{ j \in J ~|~ \top \notin A_{\star, j}  \} $
\item $ A' := A \restriction (I' \times J') $
\item $ b' := b \restriction I' $
\item $ A' \in F^{I' \times J'} $
\item $ b' \in F^{I'} $
\end{itemize}

\section{Proof idea}

We need to do the following steps in given order:
\begin{enumerate}
\item Delete all rows of $(A|b)$ where $A$ has $\bot$ or $b$ has $\top$
(they are tautologies).
\item Delete all columns of $A$ that contain $\top$
(they force respective variables to be zero).
\item If $b$ contains $\bot$ then the $(\exists x)$ part cannot be satisfied,
but $y = 0$ satisfies the other part. Stop here.
\item Assume there is no $\bot$ in $b$. Use the normal Farkas.
%\item If the $(\exists x)$ part satisfies the normal Farkas, assign
%zeros to all deleted primar variables. Our $(\exists x)$ part is satisfied.
%Why is our $(\exists y)$ part unsat?
%\item If the $(\exists y)$ part satisfies the normal Farkas, assign
%zeros to all deleted dual variables.
%If it was a row of $A$ that contained $\bot$, then the corresponding
%column of $(-A^T)$ contains $\top$.
%Thanks to $\top \cdot 0 = 0$ nothing changes.
%If it was $\top$ in $b$, nothing changes either.
%Our $(\exists y)$ part is satisfied.
%Why is our $(\exists x)$ part unsat?
Whichever solution Farkas outputs, extend it with zeros on
all deleted positions.
\end{enumerate}

\end{document}
